---
title: "HDX_proj-1_main"
author: "Benjamin DAUTRIF"
date: "2023-10-26"
output:
  html_document:
    df_print: paged
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

# The easiest way to get dplyr is to install the whole tidyverse:
install.packages("tidyverse")
# Alternatively, install just dplyr:
install.packages("dplyr")

# Install HDX package from Github :
# install.packages("remotes")
remotes::install_gitlab("dickoa/rhdx")
install.packages("RCurl")
# Loading packages :
library("rhdx")
library("RCurl")
```

# Summary : 

This script installs from the *"configuration shunk"* the *"rhdx"* package. All interogations of the HDX database relies on it. 

For more informations, see :    
https://gitlab.com/dickoa/rhdx

```{r conf, echo=TRUE}
set_rhdx_config(
  hdx_site = "prod",
#  hdx_key = NULL,
  read_only = TRUE,
#  hdx_config = NULL,
#  hdx_config_file = NULL,
#  configuration = NULL
)
```

Verifying configuration by : 

```{r , echo=TRUE}
# Configuration check :
  get_rhdx_config()
```

With result beeing : 
* <HDX Configuration> 
   HDX site: prod
   HDX site url: https://data.humdata.org
   HDX API key: *

## CALL

A simple (but evolutiv) function to call on HDX-database, searching for *"key-word"* and specifying a *"limit"* for result-list. 

```{r SEARCH MODULE, echo=TRUE}
Keyword = 'Peru'
temp = search_datasets(Keyword,rows = 100)
```

HDX_SEARCH function returns a list of HDX_datasets, regarding selection criterias. We now want to extract  informations from it selection about each dataset : 


```{r SEARCH_IN_SELECTION, echo=TRUE}
SEARCH_IN_SELECTION = function(y){
  print(unique(unlist(lapply(y,function(x){get_tags_names(x)}))))
  clef_recherche = readline("Enter TAG for selection : \n")
  mask = unlist(lapply(temp,function(x){sum(get_tags_names(x)==clef_recherche)}))
  temp_selected = y[mask]
  return(temp_selected)
}
```

```{r DOWNLOAD, echo=TRUE}
# Search in selection by TAGs :
temp_1 = SEARCH_IN_SELECTION(temp)

# Access to ressources in result : 
temp_1[[1]]$get_resources()

# Destination folder 
folder = "C:/Users/Benjamin/Documents/UN/HDX/R_HDX_proj-1/HDX_proj-1/Client_1"

# execute download :
download_resource(temp_1[[1]]$get_resources()[[1]],folder = folder)
```

```{r HDX FUNCTIONS, echo=TRUE}
# Get package for char to date conversion :
library(lubridate)

# Module to perform date-specifi search in selection :
GET_DATE <- function(selection,h=FALSE){
    # Get a selection from HDX_SEARCH funtion and get all dates.
    temp_date <- lapply(selection,function(x){get_dataset_date(x)})
    # IF option h is TRUE -> print an histogram time representation
    if(h == TRUE){
      hist(table(unlist(x)))
    }
    # Each dataset has a [from;to] pattenr. Here are two columns [Min;Max] :
    min_date_raw = unlist(lapply(temp_date,function(x){print(x[1])}))
    max_date_raw = unlist(lapply(temp_date,function(x){print(x[2])}))
    # That are binded by columns :
    temp_dates = cbind(min_date_raw,max_date_raw)
    # Convert columns to date type :
    temp_date[1] = ymd(temp_date[1])
    temp_date[2] = ymd(temp_date[2])
    
    return(temp_dates)
}

GET_UNDER_DATES = function(x){
  # Uses GET_DATE method to extract dates from datasets :
  temp_dates = GET_DATE(x)
  # print given dates for choice of upper limit :
  print(temp_dates)
  # Get clipboard entry for upper limit :
  date_recherche = readline("Enter DATE_MAX for selection : \n")
  # Production of a mask for selection subset :
  mask_date = temp_dates[,1] < date_recherche
  # Subset in return :
  return(x[mask_date])
}

GET_UPPER_DATES = function(x){
  # Uses GET_DATE method to extract dates from datasets :
  temp_dates = GET_DATE(x)
  # print given dates for choice of lesser limit :
  print(temp_dates)
  # Get clipboard entry for lesser limit :
  date_recherche = readline("Enter DATE_IN for selection : \n")
  # Production of a mask for selection subset :
  ifelse(!is.na(temp_dates[,2]),temp_dates[,2],temp_dates[,1]) > date_recherche
  # Subset in return :
  return(x[mask_date])
}
```

```{r , echo=TRUE}
# Direct pull of a dataset with <HDX Dataset>
pull_HDX <- function(ID){
  tmp = pull_dataset(identifier = ID) %>%
    get_resource(1) %>%
    read_resource()
  return(tmp)
}

res = HDX_SEARCH()
# ressource = read_resource(ressources_liste[[res]])

system2("pip install lasio")
```