--- 
title: "R_http-API_client_1"
author: "Benjamin DAUTRIF"
date: "2023-10-27"
output: pdf_document
---
# Summary : 

This script insalls from the *"configuration shunk"* the *"httr2"* package. All interogations of the HDX database relies on it. 

## For more informations, see :    
https://httr2.r-lib.org/
https://cran.r-project.org/web/packages/httr2/

## URL of the reliefWeb API :
https://apidoc.rwlabs.org/

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

# Packages installation 
# install.packages("httr2")
# install.packages("rjson")

# Load the package required to http protocole management.
library("httr2")
# Load the package required to read JSON files.
library("rjson")
```

# Variables definition :

An *"endpoint"* must be chosen to define the type of documents to search in. *"parameters"* are given in two vectors : [FIELD] and [VALUE] of the same length.      

For more information on parameters, see [API documentation]("https://apidoc.rwlabs.org/parameters")

## The client is composed of 3 functions :    

**1st** one calls for the API url and defines the *"endpoint"* :

```{r REQUEST, echo=TRUE}
req_call <- function(var_endpoint){
  endpoints = c(
    "reports",
    "disasters",
    "countries",
    "jobs",
    "training",
    "sources",
    "blog",
    "book",
    "references"
  )
  req = request("https://api.reliefweb.int/v1/")
  req = req_url_path_append(req, endpoints[var_endpoint])
  return(req)
}
```

It returns a request that can be enriched with *"parameters"* with the **2nd** function :

Beware of parentheses in multi-arguments calls for URL management sake.

```{r pressure, echo=TRUE}

# First version of the function is valuable for signle argument parameter :
param_call <- function(req,field,value){

  # Add "?" for parameters compartment delimitation :
  temp = capture.output(cat(
    req$url, '?query[value]=', sep = ""
    ))
  temp = capture.output(cat(
    temp,paste(field, ":", value, collapse = "&query[value]=", sep = ""),sep = ""
    ))
  
  req = request(temp)
  resp = req_perform(req)
  res = resp_body_json(resp)
  
  return(res)
}

# Second version of the function covers multi-arguments call and operators :
param_call_2 <- function(req,field=NULL,value=NULL,operator=NULL){

  values = c()
  for(i in seq(1,length(value))){
    if(!is.null(field[i])){
      values = append(values, capture.output(cat("query[fields][]=",field[i],sep = "")))
    }
    if(!is.null(value[i])){
      values = append(values, capture.output(cat("query[value]=",value[i],sep = "")))
    }
    if(!is.null(operator[i])){
      values = append(values, capture.output(cat("query[operator]=",operator[i],sep = "")))
    }
  }
  # Add "?" for parameters compartment delimitation :
  Kick = req$url
  temp = capture.output(
    cat(values,sep = "&")
  )
  temp = capture.output(cat(Kick,temp,sep="?"))
  
  print(temp)
  
  req = request(temp)
  resp = req_perform(req)
  # res = resp_body_json(resp)
  
  return(resp)
}
```

For more informations on *"parameters"* for this API, you may want to loog the online documentation. 

The **3rd** function *summarize* the results in a sized list : 

```{r SUMMARY,echo=TRUE}
res_summary <- function(res, field, lim = 10){
# - input : { Resultat after JSon extract [res] ; The field to print in final list [field]}
# - output : a list of fields from objects extracted by [res]

eval(parse(text = paste("
for(i in seq(1,lim)){
print(res$data[i][[1]]$fields$",field,")}",sep="")))
}
```

Example :

```{r PARAMs, echo=TRUE}
var_endpoint = 1
value = "(gaza)(flood)"
field = "title"
operator = "OR"
```

```{r CASE, echo=TRUE}
req = req_call(1)
res = param_call(req = req, field = field, value = value)
res_summary(res,field = "title")



param_call_2(req,field,value,operator)
```