--- 
title: "R_http-API_client_1"
author: "Benjamin DAUTRIF"
date: "2023-10-27"
output:
  html_document:
    df_print: paged
---

# Summary : 

This script installs from the *"configuration shunk"* the *"httr2"* package. All interogations of the HDX database relies on it. 

For more informations, see :    

https://httr2.r-lib.org/

https://cran.r-project.org/web/packages/httr2/

URL of the reliefWeb API :

https://apidoc.rwlabs.org/

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

# Packages installation 
# install.packages("httr2")
# install.packages("rjson")

# Load the package required to http protocole management.
library("httr2")
# Load the package required to read JSON files.
library("rjson")
```

# Variables definition :

An *"endpoint"* must be chosen to define the type of documents to search in. *"parameters"* are given in two vectors : [FIELD] and [VALUE] of the same length.      

For more information on parameters, see [API documentation]("https://apidoc.rwlabs.org/parameters")

## The client is composed of 3 functions :    

**1st** one calls for the API url and defines the *"endpoint"* :

```{r REQUEST, echo=TRUE}
req_call <- function(var_endpoint){
  endpoints = c(
    "reports",
    "disasters",
    "countries",
    "jobs",
    "training",
    "sources",
    "blog",
    "book"
  )
  req = request("https://api.reliefweb.int/v1/")
  req = req_url_path_append(req, endpoints[var_endpoint])
  req$url = paste0(req$url,"?appname=R_HDX_client_2")
  
  return(req)
}
```

It returns a request that can be enriched with *"parameters"* with the **2nd** function :

Beware of parentheses in multi-arguments calls for URL management seek.

```{r SUMMARY,echo=TRUE}
res_summary <- function(res, field, lim = 100){
# - input : { Resultat after JSon extract [res] ; The field to print in final list [field]}
# - output : a list of fields from objects extracted by [res]
fields = c(
    "title", # reports
    "name", # disasters
    "name", # "countries",
    "title", # "jobs",
    "title", # "training",
    "name", # "sources",
    "title", # "blog",
    "title" # "book",
  )
eval(parse(text = paste("
for(i in seq(1,lim)){
print(res$data[i][[1]]$fields$",fields[field],")}",sep="")))
}
```

```{r GRAB,echo=TRUE}
# Function destined to be the last step of the search, like a "hand on it" when GRABBING specific result :
GRAB_SEARCH = function(res) {
  # Takes a result pool to print a summary (in the "title" or "name" depending on type :
  title_list = lapply(res$data, function(x) {
    print(x$fields)
  })
  # Print the list as a summary to choose items :
  print(title_list)
  # Invite to choose :
  # - specific resource :
  y = as.numeric(readline("Choose ressource to view on web : \n"))
  # - specific field (see.....) :
  field = readline("Choose a field to view : \n")
  # - optional link to the web-page
  web_page = readline("Do you want to open web-page (Y/N) : \n")
  
  item = eval(parse(text = paste0("res$data[[", y, "]]$", field)))
  if (web_page == "Y" |
      web_page == "y") {
    web_page_url = res$data[[y]]$href
    browseURL(url = web_page_url)
  }
  
  return(item)
}
```

```{r pressure, echo=TRUE}
# Perform a request that covers multi-arguments and operators :
param_call_2 <- function(endpoint = 1, limit = NULL,field=NULL,value=NULL,date.start=NULL,date.end=NULL,operator=NULL){
  
  # First call a request on the Relief Web aPI :
  req = req_call(endpoint)
  
  # limit argument (if not given in arg. se limit in IF method) :
  if (is.null(limit) == TRUE | is.numeric(limit) == TRUE) {
    limit1 <-
      ifelse(is.null(limit) == TRUE,
             "&limit=100",
             paste0("&limit=", limit))
  } else{
    stop(
      "The argument 'limit' requires a NULL or numeric value.",
      call. = FALSE
    )
  }
  # In this version, VALUES and FIELDS are PAIRED : meaning it MUST be passed as TWO VECTORS of the same size (see example) :
  values = c()
  for(i in seq(1,length(value))){
    if(!is.null(field[i])){
      values = append(values, capture.output(cat("query[fields][]=",field[i],sep = "")))
    }
    if(!is.null(value[i])){
      values = append(values, capture.output(cat("query[value]=",value[i],sep = "")))
    }
  }
  # OPERATOR is to be given only once : 
  if(!is.null(operator)){
      operator = capture.output(cat("query[operator]=",operator,sep = ""))
  }
  
  # date.start argument
  if (is.null(date.start) == TRUE |
      is.character(date.start) == TRUE) {
    date.start1 <- ifelse(
      is.null(date.start) == TRUE,
      "",
      paste0(
        "&filter[field]=date&filter[value][from]=",
        date.start,
        "T00:00:00%2B00:00"
      )
    )
  } else{
    stop(
      "The argument 'date.start' must be set to NULL (no restriction on sample start date) or
                                requires a character string in the format 'yyyy-mm-dd'.",
      call. = FALSE
    )
  }
  
  # date.end argument
  if (is.null(date.end) == TRUE | is.character(date.end) == TRUE) {
    date.end1 <- ifelse(
      is.null(date.end) == TRUE,
      "",
      paste0(
        "&filter[field]=date&filter[value][to]=",
        date.end,
        "T00:00:00%2B00:00"
      )
    )
  } else{
    stop(
      "The argument 'date.end' must be set to NULL (no restriction on sample end date) or
                              requires a character string in the format 'yyyy-mm-dd'.",
      call. = FALSE
    )
  }
  print(date.end1)
  # get the API-URL for Relief Web and ADD LIMIT :
  kickstart = paste0(req$url,limit1)
  kickstart = paste0(kickstart,date.start1)
  kickstart = paste0(kickstart,date.end1)
  # Depending on args, ADD PARAMETERS to request :
  # - each arg MUST be separated by a "&"
  if(!is.null(values)){
    temp = capture.output(cat(values,sep = "&"))
    temp = capture.output(cat(temp,operator,sep = "&"))
    temp = capture.output(cat("&",temp,sep = ""))
    temp = capture.output(cat(kickstart,temp,sep=""))
  }else{
    temp=Kick
  }
  # Print URL for debugging :
  print(temp)
  # Perform request :
  request = request(temp)
  resp = req_perform(request)
  # Extract JSON :
  res = resp_body_json(resp)
  # Print infos : 
  print("for :")
  print(res$href)
  print("had :")
  print(res$totalCount)
  print("got :")
  print(res$count)

  # Print summary by types : 
  # - endpoint = 1 -> reports -> title
  # - endpoint = 2 -> disasters -> name
  # if(endpoint==1){
  # print(unlist(lapply(res$data,function(x){x$fields$title})))
  # }else if(endpoint==2){
  # print(unlist(lapply(res$data,function(x){x$fields$name})))
  # }
  print(res_summary(res,field = endpoint,lim = limit))
  return(res)
}
```

For more informations on *"parameters"* for this API, you may want to loog the online documentation. The **3rd** function *summarize* the results in a sized list : 

##Examples :

```{r example, echo=TRUE}
endpoint = 1
value = c("Who", "What", "Where")
field = c("title","title","title")
operator = "AND"

# Simple request :
# res = param_call_2(endpoint = endpoint,limit = 100,field = field,value = value,date.start = )
# With DATE START :
res = param_call_2(endpoint = endpoint,limit = 100,field = field,value = value,date.start = "2019-01-02")
```


```{r example_2, echo=TRUE}
endpoint = 2
value = c("storm","hurrican")
field = c("name","name")
operator = "AND"

# Simple request :
# res = param_call_2(endpoint = endpoint,limit = 100,field = field,value = value,date.start = )
# With DATE START :
res = param_call_2(endpoint = endpoint,limit = 100,field = field,value = value,date.start = "2019-01-02",operator = operator)

# res = param_call_2(endpoint = 6,limit = 10,field = c("name"),value = c("congo"))

```

```{r CASE, echo=TRUE}
# Forge request :

res = param_call_2(endpoint = 1,limit = 20, field = c("title","title"), value = c("gaza","health"),operator = "AND")

res_summary(res,field = "title")

lapply(res$data,function(x){x$fields$title})

```