--- 
title: "R http API client 1"
author: "Benjamin DAUTRIF"
date: "2023-10-27"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Summary 

This script installs from the *"configuration shunk"* the *"httr2"* package. All interogations of the HDX database relies on it. 

For more informations, see :                                    
https://httr2.r-lib.org/                                    
https://cran.r-project.org/web/packages/httr2/                                    

URL of the reliefWeb API :                                    
https://apidoc.rwlabs.org/                                    

Code for parametric search is largely inspired form "Disastr.api" ; wrapper for the UN OCHA ReliefWeb Disaster Events API. Adaptation done to adapte for other ENDPOINTS (see below, eg. "reports").


```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

# Packages installation 
# install.packages("httr2")
# install.packages("rjson")

# Load the package required to http protocole management.
library("httr2")
# Load the package required to read JSON files.
library("rjson")

# DEFINE [ENDPOINTS ; FIELDS] CORRESPONDENCE TABLE : 
# Each ENDPOINTs have it's own structure in the Relief Web API 
# (see "https://apidoc.reliefweb.int/fields-tables" for more).
# For this to be transparent to user, function needs to switch those on demande.
# RECALL : ENDPOINTs are to be chosen in a list, same as following :
endpoints = c(
    "reports",
    "disasters",
    "countries",
    "jobs",
    "training",
    "sources",
    "blog",
    "book"
  )
   
# Thus the corresponding fields-table for TITLE/NAME :  
fields = c(
    "title", # reports
    "name",  # disasters
    "name",  # "countries",
    "title", # "jobs",
    "title", # "training",
    "name",  # "sources",
    "title", # "blog",
    "title"  # "book",
  )

Param_table = cbind(fields,endpoints)
```

# Variables definition :

*"Endpoint"* defines the type of documents (eg, reports, job annonce, disaster event....). *"parameters"* are then given to search for words in specific fields ; some being specific to certain *endpoints*.      

For more information on parameters, see [API documentation]("https://apidoc.rwlabs.org/parameters")

# Client 

3 functions :    

**1.** : calls API url and defines *"endpoint"*

```{r REQUESTING, echo=TRUE}
req_call <- function(endpoint){
  
  req = request("https://api.reliefweb.int/v1/")
  req = req_url_path_append(req, endpoints[endpoint])
  req$url = paste0(req$url,"?appname=R_HDX_client_2")
  
  return(req)
}
```

Returns a REQUEST which is to enrich with *"parameters"*. 

**2.** Summaryze results :

Beware of parentheses in multi-arguments calls for URL management seek.

```{r SUMMARYING,echo=TRUE}
res_summary <- function(res, type){

  # - input : { Resultat after JSon extract [res] ; The field to print in final list [field]}
  # - output : a list of fields from objects extracted by [res]
  
  # SUMMARY (depending on types as ENDPOINT, due to different API structures ; see API) : 
    # Example :
    # - endpoint = 1 -> reports -> title
    # - endpoint = 2 -> disasters -> name
    
x = Param_table[,2] == type
field = Param_table[x,1]

eval(parse(text = paste("
for(i in seq(1,res$count)){
print(paste0(i,\" -> \",res$data[i][[1]]$fields$",field,"))}",sep="")))
}
```

**3.** Get specific FIELD in specific resource and see on web :

```{r COLLECTING,echo=TRUE}
COLLECT_SEARCH = function(res) {
  # Objectif : print a summary according to given FIELD
  # Input : a result-pool 
  # Output : LIST of fields for all pool
  # PRINT MENUE :
  title_list = lapply(res$data, function(x) {
    print(x$fields)
  })
  print(title_list)
  # KEYBOARD ENTRY :
  # - CHOOSE resource :
  y = as.numeric(readline("Choose ressource to view on web : \n"))
  # - CHOOSE field (see.....) :
  field = readline("Choose a field to view : \n")
  # - OPTIONAL web-page :
  web_page = readline("Do you want to open web-page (Y/N) : \n")
  # LIST FIELDS in the pool :
  result = eval(parse(text = paste0("res$data[[", y, "]]$", field)))
  # WEB PAGE :
  if (web_page == "Y" |
      web_page == "y") {
    web_page_url = res$data[[y]]$href
    browseURL(url = web_page_url)
  }
  # RETURN the field-list 
  return(result)
}
```

```{r PRINTING,echo=TRUE}
PRINT_SEARCH = function(res,
                        numero,
                        namefile,
                        endpoint,
                        txt = NULL,
                        html = NULL,
                        pdf = NULL) {

  # Get specific ENDPOINT item :
  x = Param_table[,2] == endpoints[endpoint]
  field = Param_table[x,1]

  # Menu results : 
  lapply(res$data,function(x){x$fields$name})
  
  # Request URL :
  url  = res[numero,"data.href"]
  req  = request(url)
  resp = req_perform(req)
  resp = resp_body_json(resp)
  
  # Print resource to txt file :
  if(!is.null(txt)){
  # sink(
  #   file = paste0(
  #     "C:/Users/Benjamin/Documents/GITHub/HDX_Proj-1/DATAs/Logs/",
  #     namefile,
  #     ".txt"
  #   ),
  #   append = FALSE,
  #   type = c("output")
  # )
  resultat.txt =  ifelse(endpoint == 1,
                    resp$data[[1]]$fields$body,
                  ifelse(endpoint == 2,
                    resp$data[[1]]$fields$description,
                  "c'est autre chose"
  ))
  
  sink(
    file = paste0(
      "C:/Users/Benjamin/Documents/GITHub/HDX_Proj-1/DATAs/TXT/",
      namefile,
      ".txt"
    ))
    print(resultat.txt)
  sink()
  }
  
 # Print resource as HTML : 
  if(!is.null(html)){

  resultat.html =  ifelse(endpoint == 1,
                      resp$data[[1]]$fields$`body-html`,
                   ifelse(endpoint == 2,
                      resp$data[[1]]$fields$`description-html`,
                      "c'est autre chose"
  ))
  sink(
    file = paste0(
      "C:/Users/Benjamin/Documents/GITHub/HDX_Proj-1/DATAs/HTML/",
      namefile,
      ".html"
    ))
    print(resultat.html)
  sink()
  }
  
  # Save pdf file :
  if (!is.null(pdf)) {
  # Specify the URL of the file you want to download
  url <- resp$data[[1]]$fields$file[[1]]$url
  
  # Specify the file name and location where you want to save the file on your computer
  file_path <- "C:/Users/Benjamin/Documents/GITHub/HDX_Proj-1/DATAs/PDF/"
  
  # Call the download.file() function, passing in the URL and file name/location as arguments
  download.file(url, paste(file_path, namefile,".pdf" , sep = ""), mode = "wb")
  }    
    # sink(file = paste0("C:/Users/Benjamin/Documents/GITHub/HDX_Proj-1/Logs/",namefile,"_pdf.txt"), append = FALSE, type = c("output"))
    # if(pdf==TRUE){
    # library(pdftools)
    # pdftools::pdf_text(pdf = req$body)
    # }
    # sink()
  
  return(resp_body_json(req_perform(req)))
}
```

Main function :  

```{r MAIN, echo=TRUE}
# Perform a request that covers multi-arguments and operators :
Main <-
  function(endpoint = NULL,
           limit = NULL,
           operator = NULL,
           date.start = NULL,
           date.end = NULL,
           ascending.date = NULL,
           ascending.title = NULL,
           exact.match = FALSE,
           country = NULL,
           disaster = NULL,
           field = NULL,
           value = NULL
           ) {
    
  # DEBUG panel : exposes variables to run function from the inside. Uncomment if needed.
  # - Set intern variables : 
            # endpoint = 1
            # limit = 50
            # field = c("title", "country")
            # value = c("water","france")
            # date.start = "2019-01-02"
            # date.end = "2020-01-02"
            # ascending.date = TRUE
            # ascending.title = TRUE
            # exact.match = FALSE
            # operator = NULL
            # disaster = NULL
            # country = NULL

   # 1. CALL REQUEST ON THE RELIEF WEB API :
   req = req_call(endpoint)
   
   # 2. PARAMETRISATION : 
   # Conversation with API goes through URLs. First strep is to create each parameter code.
   # - LIMIT argument (if not given in arg. se limit in IF method) :
   if (is.null(limit) == TRUE | is.numeric(limit) == TRUE) {
     limit1 <-
       ifelse(is.null(limit) == TRUE,
              "&limit=100",
              paste0("&limit=", limit))
   }
   
   # - OPERATOR (given only once) :
   if (!is.null(operator)) {
     operator1 = capture.output(cat("query[operator]=", operator, sep = ""))
   }
   
   # - DATE.START argument
   if (is.null(date.start) == TRUE |
       is.character(date.start) == TRUE) {
     date.start1 <- ifelse(
       is.null(date.start) == TRUE,
       "",
       paste0(
         "&filter[field]=date&filter[value][from]=",
         date.start,
         "T00:00:00%2B00:00"))
   }
   
   # - DATE.END argument
   if (is.null(date.end) == TRUE | is.character(date.end) == TRUE) {
     date.end1 <- ifelse(
       is.null(date.end) == TRUE,
       "",
       paste0(
         "&filter[field]=date&filter[value][to]=",
         date.end,
         "T00:00:00%2B00:00"))
   }
   
   # - ASCENDING DATE argument
   if (is.logical(ascending.date) == TRUE) {
     ascending1 <-
       ifelse(ascending.date == FALSE, "&sort[]=date:desc", "&sort[]=date:asc")
   }
   
   # - ASCENDING TITLE argument
   if (is.logical(ascending.title) == TRUE) {
     ascending1 <-
       ifelse(
         ascending.title == FALSE,
         paste0("&sort[]=", fields[endpoint], ":desc"),
         paste0("&sort[]=", fields[endpoint], ":asc")
       )
   }
   
   # - Exact.match argument
   if ((
     is.character(disaster) == TRUE | is.character(country) == TRUE )) {
     if (is.logical(exact.match) == TRUE) {
       exact.match1 <- ifelse(exact.match == FALSE,
                              "&query[operator]=OR",
                              "&query[operator]=AND")
     }
   }
   
   # - Country argument :
   if (is.null(country) == TRUE | is.character(country) == TRUE) {
     country1 <- ifelse(is.null(country) == TRUE,
                        "",
                        paste0("&filter[field]=country:", paste(gsub(
                          "\\s{1}", "%", country
                        ), collapse = "|")))
   }
   
    # - Disaster argument
   if (is.null(disaster) == TRUE | is.character(disaster) == TRUE) {
     disaster1 <- ifelse(is.null(disaster) == TRUE,
                         "",
                         paste0("&query[value]=type:", paste(
                           gsub("\\s{1}", "%", disaster), collapse = "|")))
   }
   
   # - FORMATING  (query) ARGUMENTS :
   # CAUTION : In this version, VALUES and FIELDS are PAIRED :  MUST be  TWO VECTORS of the SAME SIZE :
   # EXAMPLE : to search for two words in the same field (here "title"), see
   # Main(...,field = c("title, "title"),value = c("word_1", "word_2"), )
   # Debug panel : 
     # value = c("gaza","health")
     # field = c("title","title")
     # operator = c("AND")
  
   field1 = c()
   value1 = c()
   operator1 = c()
   for (i in seq(1, length(value))) {
     if (!is.null(field[i])) {
       field1 = append(field1, capture.output(cat("query[fields][]=", field[i], sep = "")))
     }
     if (!is.null(value[i])) {
       value1 = append(value1, capture.output(cat("query[value]=", value[i], sep = "")))
     }
   }
   if (!is.null(operator)) {
       operator1 = capture.output(cat("query[operator]=", operator, sep = ""))
     }
   
  # 4. Add PARAMETEERS to RELIEF WEB - API-URL :
  # Makes a "Kickstart" with individual concatenated parametric codes.
  # - Each line verifying argument existence before concatenation.
    kickstart = paste0(limit1)
    kickstart = ifelse(exists("ascending1"),capture.output(cat(kickstart,ascending1,sep = "")),kickstart)
    kickstart = ifelse(exists("country1"),capture.output(cat(kickstart,country1,sep = "")),kickstart)
    kickstart = ifelse(exists("date.start1"),capture.output(cat(kickstart,date.start1,sep = "")),kickstart)
    kickstart = ifelse(exists("date.end1"),capture.output(cat(kickstart,date.end1,sep = "")),kickstart)
    kickstart = ifelse(exists("disaster1"),capture.output(cat(kickstart,disaster1,sep = "")),kickstart)
    kickstart = ifelse(exists("exact.match1"),capture.output(cat(kickstart,country1,sep = "")),kickstart)

  # - Then, concatenation with [API-url + endpoint-code] :
    kickstart = paste0(req$url,kickstart)
    
  # - Then qdd to the KICKSTART the code for every [KEY WORDS ; FIELDS] couples. 
  # RECALL : words and fields a to give by couple. Have to be presented as equaly sized VECTOR
  if(!is.null(field1) & !is.null(value1)){
    # vectors are of the form : 
    # ["query[fields][]=.........." ; ""query[value]=.........."]
    # CAUTION : "fields" and "values" are PAIRED, so their vectors MUST have same length
    temp_params = paste(field1,value1,sep = "&")
    temp_params = capture.output(cat(temp_params,sep = "&"))
    # Operator (for now) is just to be passed once
    temp =        capture.output(cat(temp_params,operator1,sep = "&"))

    # temp = capture.output(cat(temp,operator,sep = "&"))
    # arranged in the form : [.................]
    # - Then concatenate QUERY-parametric code to KICKSTART
    # -- at first add a API specific symbole "&" for arguments concatenation
    temp = capture.output(cat("&",temp,sep = ""))
    # -- then concatenate KICKSTART with ARGUMENTS
    temp = capture.output(cat(kickstart,temp,sep=""))
  }else{
    # - If no QUERY-argument, let the KICKSTART be
    temp=kickstart
  }
    
  # Print URL :
  print(temp)
  
  # PERFORM REQUEST :
  request = request(temp)
  resp = req_perform(request)
  
  # JSON :
  res = resp_body_json(resp, simplifyVector = TRUE)
  res = as.data.frame(res)
  # INFOS : 
  print("for :")
  print(res$href[1])
  print("had :")
  print(res$totalCount[1])
  print("got :")
  print(res$count[1])

  # SUMMARY (depending on types as ENDPOINT, due to different API structures ; see API) : 
  # Example :
  # - endpoint = 1 -> reports -> title
  # - endpoint = 2 -> disasters -> name
  print(str(x,max.level = 2,vec.len = 0))
  eval(parse(text = paste("print(res$data.fields$",fields[endpoint],")",sep = "")))
  # Return list of results :
  return(res)
}
```

For more informations on *"parameters"* for this API, you may want to loog the online documentation. The **3rd** function *summarize* the results in a sized list : 

## Examples :

```{r EXAMPLE 1, echo=TRUE}
endpoint = 2
value = c("flood")
field = c("name")
operator = "OR"

# Simple request :
# res = Main(endpoint = endpoint,limit = 100,field = field,value = value,date.start = )
# With DATE START :
res = Main(endpoint = endpoint,limit = 20, ascending.date = TRUE, field = field, value = value, operator = operator)
```


```{r EXAMPLE 2, echo=TRUE}
endpoint = 2
value = c("cuba","storm")
field = c("name","name")
# operator = "OR"

# Simple request :
# res = Main(endpoint = endpoint,limit = 100,field = field,value = value,date.start = )
# With DATE START :
res = Main(
  endpoint = endpoint,
  limit = 100,
  field = field,
  value = value,
  ascending.date = FALSE,
  operator = "OR"
)

```

```{r CASE, echo=TRUE}
# Forge request :
# res = Main(endpoint = 1,limit = 20, field = c("title","title"), value = c("gaza","health"),operator = "AND")

endpoint = 1
limit = 100
value = c("WHO", "congo")
# RECALL : field must correspond to result structure (see Relief Web API) :
# Either :
field = c("title", "title")
# Either :
# field = c("name")
# Paramatrisation :
# operator = "AND"
date.start = "2020-01-02"
date.end = "2020-01-02"
namefile = "test_3"

# Simple request parameters :
# res = Main(endpoint = endpoint,limit = 100,field = field,value = value,date.start = )
res = Main(
  endpoint = endpoint,
  limit = limit,
  field = field,
  value = value,
  # disaster = "storm",
  date.start = date.start
)

numero = 14

test_1 = PRINT_SEARCH(
  res = res,
  numero = numero,
  endpoint = endpoint,
  namefile = namefile,
  txt = TRUE,
  html = TRUE
)

# Request collections :

# Search for epidemiologic events for DENGUE disease, chronologicaly ordered :
# x = Main(endpoint = 2,ascending.date = FALSE,disaster = "Epidemic", value = c("Dengue"),field = c("name"))
# res_summary(x,type = "disasters")

# Search for latest REPORTS with WHO in their title :
# x = Main(endpoint = 1,limit = 50,ascending.date = FALSE, value = c("WHO"),field = c("title"))
# res_summary(x,type = "reports")

```

```{r , echo=TRUE}
res = Main(endpoint = 2,limit = 20,date.start = "2015-01-02",ascending.date = FALSE,value = c("Epidemic"),field = c("name"))

```